### 2025-06-12 cross_encoder/thesis/Appendices/AppendixA.tex
Refactor results presentation and update appendix in thesis
- Changed table format in Chapter 5 to include epoch-wise results for models.
- Enhanced clarity of evaluation results by adjusting column headers and adding a footnote for metric definitions.
- Updated the table layout to a two-column format for better readability.
- Revised the table of contents to reflect changes in section numbering due to added content.
- Added detailed evaluation metrics in Appendix A, including primary ranking metrics, precision and recall analysis, NDCG analysis, and metric definitions.
- Adjusted references a

\chapter{Detailed Evaluation Metrics} % Main appendix title

\label{AppendixA}

This appendix presents comprehensive evaluation metrics for all model configurations tested in our experiments. While the main chapters focus on key performance indicators, these detailed tables provide additional insights into model behavior across various retrieval effectiveness measures.

\section{Primary Ranking Metrics}
Table~\ref{tab:primary_metrics} presents the main effectiveness metrics including MAP, NDCG@10, MRR@10, and Precision@10 for all tested configurations.

\begin{table}[htbp]
\centering
\caption{Primary ranking metrics comparing cross-encoder models across training epochs.}
\label{tab:primary_metrics}
\small
\begin{tabular}{llccccc}
\toprule
\textbf{Model} & \textbf{Config} & \textbf{MAP} & \textbf{NDCG@10} & \textbf{MRR@10} & \textbf{P@10} & \textbf{R-Prec} \\
\midrule
\multirow{6}{*}{GTE} & AdamW-1 & 0.5005 & 0.7224 & 0.5940 & 0.8116 & 0.4964 \\
 & AdamW-2 & 0.4999 & 0.7203 & 0.5942 & 0.8256 & 0.5018 \\
 & AdamW-3 & 0.4899 & 0.6902 & 0.5972 & 0.8093 & 0.5017 \\
 & Lion-1 & 0.4794 & 0.6970 & 0.5854 & 0.7884 & 0.4814 \\
 & Lion-2 & 0.4577 & 0.6792 & 0.5957 & 0.7721 & 0.4642 \\
 & Lion-3 & 0.4521 & 0.6571 & 0.5931 & 0.7488 & 0.4662 \\
\midrule
\multirow{6}{*}{MiniLM} & AdamW-1 & 0.4814 & 0.7008 & 0.5828 & 0.8000 & 0.4884 \\
 & AdamW-2 & 0.4891 & 0.7094 & 0.5818 & 0.8116 & 0.4916 \\
 & AdamW-3 & 0.4908 & 0.7127 & 0.5826 & 0.8116 & 0.4943 \\
 & Lion-1 & 0.4858 & 0.7031 & 0.5890 & 0.8140 & 0.4952 \\
 & Lion-2 & 0.4755 & 0.6916 & 0.5942 & 0.8070 & 0.4803 \\
 & Lion-3 & 0.4706 & 0.6808 & \textbf{0.5988} & 0.8070 & 0.4809 \\
\midrule
\multirow{6}{*}{ModernBERT} & AdamW-1 & 0.5066 & 0.7105 & 0.5866 & 0.8163 & 0.5161 \\
 & AdamW-2 & 0.4893 & 0.6839 & 0.5886 & 0.7814 & 0.4946 \\
 & AdamW-3 & 0.4971 & 0.6959 & 0.5916 & 0.7860 & 0.5116 \\
 & Lion-1 & \textbf{0.5121} & 0.7142 & 0.5834 & 0.8163 & 0.5148 \\
 & Lion-2 & 0.5115 & \textbf{0.7225} & 0.5908 & 0.8209 & \textbf{0.5183} \\
 & Lion-3 & 0.5020 & 0.7051 & \textbf{0.5988} & \textbf{0.8256} & 0.5102 \\
\bottomrule
\end{tabular}
\end{table}

\section{Precision and Recall Analysis}
Tables~\ref{tab:precision_metrics} and~\ref{tab:recall_metrics} provide detailed precision and recall values at different cutoff thresholds, revealing the models' behavior at various result list depths.

\begin{table}[htbp]
\centering
\caption{Precision metrics at different cutoff levels.}
\label{tab:precision_metrics}
\small
\begin{tabular}{llcccc}
\toprule
\textbf{Model} & \textbf{Config} & \textbf{P@5} & \textbf{P@10} & \textbf{P@20} & \textbf{P@100} \\
\midrule
\multirow{6}{*}{GTE} & AdamW-1 & 0.8930 & 0.8116 & 0.7372 & 0.3965 \\
 & AdamW-2 & 0.8698 & 0.8256 & 0.7326 & 0.3963 \\
 & AdamW-3 & 0.8558 & 0.8093 & 0.7128 & 0.3942 \\
 & Lion-1 & 0.8558 & 0.7884 & 0.7070 & 0.3893 \\
 & Lion-2 & 0.8326 & 0.7721 & 0.6802 & 0.3740 \\
 & Lion-3 & 0.8140 & 0.7488 & 0.6907 & 0.3740 \\
\midrule
\multirow{6}{*}{MiniLM} & AdamW-1 & 0.8698 & 0.8000 & 0.7151 & 0.3870 \\
 & AdamW-2 & 0.8698 & 0.8116 & 0.7186 & 0.3895 \\
 & AdamW-3 & 0.8558 & 0.8116 & 0.7256 & 0.3891 \\
 & Lion-1 & 0.8651 & 0.8140 & 0.7233 & 0.3847 \\
 & Lion-2 & 0.8744 & 0.8070 & 0.7081 & 0.3821 \\
 & Lion-3 & 0.8465 & 0.8070 & 0.6930 & 0.3809 \\
\midrule
\multirow{6}{*}{ModernBERT} & AdamW-1 & 0.8651 & 0.8163 & 0.7349 & 0.3993 \\
 & AdamW-2 & 0.8605 & 0.7814 & 0.6953 & 0.3912 \\
 & AdamW-3 & 0.8465 & 0.7860 & 0.6988 & 0.3958 \\
 & Lion-1 & 0.8791 & 0.8163 & 0.7314 & 0.4033 \\
 & Lion-2 & 0.8698 & \textbf{0.8209} & 0.7279 & \textbf{0.4016} \\
 & Lion-3 & 0.8651 & 0.8256 & 0.7093 & 0.3967 \\
\bottomrule
\end{tabular}
\end{table}

\begin{table}[htbp]
\centering
\caption{Recall metrics at different cutoff levels.}
\label{tab:recall_metrics}
\small
\begin{tabular}{llcccc}
\toprule
\textbf{Model} & \textbf{Config} & \textbf{Recall@5} & \textbf{Recall@10} & \textbf{Recall@20} & \textbf{Recall@100} \\
\midrule
\multirow{6}{*}{GTE} & AdamW-1 & 0.1051 & 0.1689 & 0.2815 & 0.5501 \\
 & AdamW-2 & 0.1041 & 0.1760 & 0.2753 & 0.5538 \\
 & AdamW-3 & 0.1051 & 0.1715 & 0.2692 & 0.5489 \\
 & Lion-1 & 0.1002 & 0.1614 & 0.2684 & 0.5401 \\
 & Lion-2 & 0.0973 & 0.1610 & 0.2545 & 0.5211 \\
 & Lion-3 & 0.0929 & 0.1542 & 0.2574 & 0.5181 \\
\midrule
\multirow{6}{*}{MiniLM} & AdamW-1 & 0.1027 & 0.1652 & 0.2645 & 0.5426 \\
 & AdamW-2 & 0.1013 & 0.1668 & 0.2676 & 0.5446 \\
 & AdamW-3 & 0.1007 & 0.1702 & 0.2719 & 0.5427 \\
 & Lion-1 & 0.1022 & 0.1688 & 0.2704 & 0.5338 \\
 & Lion-2 & 0.1025 & 0.1641 & 0.2613 & 0.5338 \\
 & Lion-3 & 0.1001 & 0.1676 & 0.2542 & 0.5249 \\
\midrule
\multirow{6}{*}{ModernBERT} & AdamW-1 & 0.0995 & 0.1678 & 0.2731 & 0.5520 \\
 & AdamW-2 & 0.0979 & 0.1634 & 0.2603 & 0.5443 \\
 & AdamW-3 & 0.1019 & 0.1623 & 0.2627 & 0.5476 \\
 & Lion-1 & 0.1052 & 0.1689 & 0.2733 & 0.5542 \\
 & Lion-2 & 0.1048 & \textbf{0.1732} & 0.2699 & \textbf{0.5608} \\
 & Lion-3 & 0.1037 & 0.1722 & 0.2640 & 0.5570 \\
\bottomrule
\end{tabular}
\end{table}

\section{NDCG Analysis}
Table~\ref{tab:ndcg_metrics} provides normalized discounted cumulative gain values at multiple cutoff points.

\begin{table}[htbp]
\centering
\caption{NDCG metrics at different cutoff levels.}
\label{tab:ndcg_metrics}
\small
\begin{tabular}{llcccc}
\toprule
\textbf{Model} & \textbf{Config} & \textbf{NDCG@5} & \textbf{NDCG@10} & \textbf{NDCG@20} & \textbf{NDCG@100} \\
\midrule
\multirow{6}{*}{GTE} & AdamW-1 & 0.7567 & 0.7224 & 0.7071 & 0.6591 \\
 & AdamW-2 & 0.7343 & 0.7203 & 0.6967 & 0.6550 \\
 & AdamW-3 & 0.7080 & 0.6902 & 0.6703 & 0.6426 \\
 & Lion-1 & 0.7229 & 0.6970 & 0.6786 & 0.6433 \\
 & Lion-2 & 0.7008 & 0.6792 & 0.6580 & 0.6248 \\
 & Lion-3 & 0.6801 & 0.6571 & 0.6450 & 0.6124 \\
\midrule
\multirow{6}{*}{MiniLM} & AdamW-1 & 0.7207 & 0.7008 & 0.6833 & 0.6402 \\
 & AdamW-2 & 0.7345 & 0.7094 & 0.6908 & 0.6471 \\
 & AdamW-3 & 0.7244 & 0.7127 & 0.6949 & 0.6474 \\
 & Lion-1 & 0.7163 & 0.7031 & 0.6829 & 0.6369 \\
 & Lion-2 & 0.7099 & 0.6916 & 0.6706 & 0.6295 \\
 & Lion-3 & 0.6837 & 0.6808 & 0.6584 & 0.6268 \\
\midrule
\multirow{6}{*}{ModernBERT} & AdamW-1 & 0.7171 & 0.7105 & 0.6936 & 0.6555 \\
 & AdamW-2 & 0.7112 & 0.6839 & 0.6701 & 0.6431 \\
 & AdamW-3 & 0.7156 & 0.6959 & 0.6729 & 0.6497 \\
 & Lion-1 & 0.7285 & 0.7142 & 0.6991 & 0.6610 \\
 & Lion-2 & \textbf{0.7381} & \textbf{0.7225} & \textbf{0.6978} & \textbf{0.6638} \\
 & Lion-3 & 0.7165 & 0.7051 & 0.6797 & 0.6542 \\
\bottomrule
\end{tabular}
\end{table}

\section{Metric Definitions}
\label{app:metric_defs}
Table~\ref{tab:metric_definitions} provides formal definitions of all evaluation metrics used in our experimental analysis.

\begin{table}[htbp]
\centering
\caption{Definitions of key evaluation metrics used in this thesis.}
\label{tab:metric_definitions}
\small
\begin{tabularx}{\textwidth}{lX}
\toprule
\textbf{Metric} & \textbf{Definition} \\
\midrule
MAP & Mean Average Precision. The mean of average precision scores across all queries, where average precision is the average of precision values calculated at every position where a relevant document is retrieved. Range: [0,1]. \\
NDCG@k & Normalized Discounted Cumulative Gain at rank k. Measures ranking quality with graded relevance and position-based discount. Range: [0,1]. \\
P@k & Precision at k. The proportion of retrieved documents in the top k results that are relevant. Range: [0,1]. \\
Recall@k & Recall at k. The proportion of all relevant documents that are retrieved in the top k results. Range: [0,1]. \\
MRR@10 & Mean Reciprocal Rank at 10. Average of the reciprocal ranks of the first relevant document within the top 10 results. Range: [0,1]. \\
R-Prec & R-Precision. Precision after R documents retrieved, where R is the number of relevant documents for the query. Range: [0,1]. \\
bpref & Binary Preference. Measures ranking quality when relevance judgments are incomplete by considering known relevant/non-relevant document pairs. Range: [0,1]. \\
\bottomrule
\end{tabularx}
\end{table}