### 2025-06-13 cross_encoder/thesis/Chapters/3.research_gap_and_proposed_solution.tex

% Chapter Template

\chapter{Research Gap \& Methodology} % Main chapter title

\label{Chapter3} % Change X to a consecutive number; for referencing this chapter elsewhere, use \ref{Chapter3}

\section{Research Gap}

Despite the established effectiveness of cross-encoder models for information retrieval, several critical gaps exist in current research that motivate this investigation:

\subsection{Limited Optimization Strategy Exploration}

While AdamW has become the de facto standard for transformer training, the systematic investigation of optimization strategies for cross-encoder models remains limited. Most studies assume AdamW without exploring alternatives, particularly for information retrieval tasks where ranking precision is critical. The recently proposed Lion optimizer \cite{chen2023symbolic}, with its novel sign-based momentum approach, presents an unexplored opportunity for cross-encoder optimization.

\subsection{Architecture-Optimizer Interaction Analysis}

Different transformer architectures may respond differently to optimization strategies, yet this interaction has not been systematically studied. Modern architectures like:

\begin{itemize}
    \item \textbf{ModernBERT} \cite{modernbert}: Incorporating Rotary Positional Embeddings (RoPE), Flash Attention, and GeGLU activations with 8192 token context support
    \item \textbf{GTE models} \cite{li2023towards}: Multilingual architectures trained on diverse corpora with extended context capabilities
    \item \textbf{MiniLM} \cite{wang2020minilm}: Efficient distilled variants designed for computational efficiency
\end{itemize}

Each may exhibit different optimization dynamics with various optimizers, requiring systematic evaluation.

\subsection{Information Retrieval Context Specificity}

Most optimizer comparisons focus on language modeling or general NLP tasks. Information retrieval presents unique challenges:
\begin{itemize}
    \item \textbf{Ranking Quality Focus}: IR metrics prioritize ranking precision over classification accuracy
    \item \textbf{Query-Document Interaction}: Cross-encoder architectures require different optimization considerations than standard sequence models
    \item \textbf{Training Efficiency}: Large-scale IR datasets demand memory-efficient training approaches
\end{itemize}

\subsection{Modern Transformer Architecture Gaps}

Recent architectural innovations have not been thoroughly evaluated with alternative optimizers:

\subsubsection{ModernBERT Architectural Advances}
ModernBERT \cite{modernbert} represents the current state-of-the-art in encoder-only transformers, incorporating several key innovations:

\begin{itemize}
    \item \textbf{Rotary Positional Embeddings (RoPE)} \cite{su2023roformerenhancedtransformerrotary}: Enable effective handling of longer sequences up to 8192 tokens by providing more flexible positional encoding compared to traditional absolute positional embeddings.
    \item \textbf{Flash Attention} \cite{dao2022flashattentionfastmemoryefficientexact}: Improves memory efficiency and computational speed for attention operations, particularly beneficial for longer sequences.
    \item \textbf{GeGLU Activation Functions} \cite{shazeer2020gluvariantsimprovetransformer}: Enhance model expressiveness while maintaining computational efficiency compared to traditional ReLU activations.
\end{itemize}

These architectural improvements enable efficient processing of long documents but their interaction with different optimizers, particularly Lion's memory-efficient approach, remains unexplored.

\subsubsection{GTE Model Family}
The General Text Embeddings (GTE) family \cite{li2023towards} demonstrated that models trained on diverse, multilingual corpora could achieve strong performance across various text understanding tasks. The GTE-multilingual-base model supports 75 languages and 8192 token context length, but optimization strategy comparisons for this architecture in IR contexts are lacking.

\subsection{Comprehensive IR Benchmarking Gap}

Existing optimizer evaluations typically focus on limited metrics or datasets. A comprehensive evaluation across standard IR benchmarks (MS MARCO, TREC DL) with multiple effectiveness metrics (NDCG@10, MAP, MRR@10, Recall@10, R-Prec, P@10) is needed to establish optimizer performance in IR contexts.

\subsection{Training Efficiency and Memory Usage Analysis}

Lion optimizer claims memory efficiency advantages due to its simplified momentum tracking, but these benefits have not been validated for computationally intensive cross-encoder training scenarios typical in information retrieval applications.

\section{Methodology}

This research addresses the identified gaps through a systematic comparative study of Lion and AdamW optimizers across three distinct cross-encoder architectures for information retrieval tasks.

\subsection{Research Objectives}

\begin{itemize}
    \item \textbf{Primary Objective}: Compare Lion and AdamW optimizer effectiveness for cross-encoder reranking across multiple transformer architectures
    \item \textbf{Secondary Objectives}: 
    \begin{itemize}
        \item Analyze architecture-optimizer interaction patterns
        \item Evaluate training dynamics and convergence behavior
        \item Assess memory efficiency and computational requirements
        \item Provide comprehensive IR benchmark results
    \end{itemize}
\end{itemize}

\subsection{Model Selection Rationale}

Three transformer architectures representing different efficiency-effectiveness trade-offs:

\subsubsection{MiniLM-L12-H384-uncased}
\begin{itemize}
    \item \textbf{Architecture}: Compact model (384 hidden dimensions, 12 layers)
    \item \textbf{Design Focus}: Efficiency through knowledge distillation from larger BERT models
    \item \textbf{Context Length}: Standard 512 tokens
    \item \textbf{Use Case}: Resource-constrained environments requiring fast inference
\end{itemize}

\subsubsection{GTE-multilingual-base}
\begin{itemize}
    \item \textbf{Architecture}: Multilingual encoder with transformer++ backbone (BERT + RoPE + GLU)
    \item \textbf{Context Support}: 8192 tokens for long document processing
    \item \textbf{Language Coverage}: 75+ languages with strong cross-lingual capabilities
    \item \textbf{Training Data}: Diverse multilingual corpora for robust text understanding
\end{itemize}

\subsubsection{ModernBERT-base}
\begin{itemize}
    \item \textbf{Architecture}: State-of-the-art encoder with advanced optimizations
    \item \textbf{Key Features}: RoPE, Flash Attention, GeGLU activations
    \item \textbf{Context Length}: 8192 tokens with efficient long-context processing
    \item \textbf{Training Scale}: 2 trillion tokens for comprehensive language understanding
\end{itemize}

\subsection{Experimental Design}

\subsubsection{Training Configuration}

\textbf{Infrastructure}: Modal cloud platform with NVIDIA A100-80GB GPUs providing consistent computational environment

\textbf{Training Parameters}:
\begin{itemize}
    \item \textbf{Batch Size}: 16 per GPU for memory efficiency
    \item \textbf{Training Duration}: 3 epochs with checkpoint evaluation after each epoch
    \item \textbf{Dataset}: MS MARCO passage ranking triplets (~2M query-passage pairs)
    \item \textbf{Precision}: BF16 for improved training efficiency
    \item \textbf{Random Seed}: 12 for reproducible results
\end{itemize}

\subsubsection{Optimizer Configuration Details}

\textbf{AdamW Configuration}:
\begin{itemize}
    \item \textbf{Learning Rates}: 
    \begin{itemize}
        \item 2e-5 for MiniLM and GTE (standard rate)
        \item 2e-5 for ModernBERT (baseline comparison)
    \end{itemize}
    \item \textbf{Weight Decay}: 0.01 for regularization
    \item \textbf{Beta Parameters}: β₁=0.9, β₂=0.999 (default values)
    \item \textbf{Scheduler}: Linear warmup for 10\% of training steps
\end{itemize}

\textbf{Lion Configuration}:
\begin{itemize}
    \item \textbf{Learning Rates}: 
    \begin{itemize}
        \item 2e-5 for MiniLM and GTE (consistency with AdamW)
        \item 2e-6 for ModernBERT (optimized based on architecture characteristics)
    \end{itemize}
    \item \textbf{Beta Parameters}: β₁=0.9, β₂=0.99 (recommended values)
    \item \textbf{Weight Decay}: 0.01 for fair comparison
    \item \textbf{Scheduler}: Cosine Annealing for ModernBERT, linear for others
\end{itemize}

\subsection{Evaluation Protocol}

\subsubsection{Datasets and Metrics}

\textbf{Primary Evaluation}: TREC 2019 Deep Learning Track passage ranking
\begin{itemize}
    \item \textbf{Query Set}: 43 queries with graded relevance judgments
    \item \textbf{Candidate Retrieval}: Top-1000 BM25 candidates per query
    \item \textbf{Metrics}: NDCG@10, MAP, Recall@10, R-Precision, P@10
\end{itemize}

\textbf{Secondary Evaluation}: MS MARCO passage development set
\begin{itemize}
    \item \textbf{Focus Metric}: MRR@10 (Mean Reciprocal Rank at cutoff 10)
    \item \textbf{Purpose}: Complementary perspective on larger query set
\end{itemize}

\subsubsection{Evaluation Framework}

\begin{itemize}
    \item \textbf{Reranking Process}: Cross-encoder scoring of BM25 candidate sets
    \item \textbf{Evaluation Tools}: Standard trec\_eval for consistent metric calculation
    \item \textbf{Checkpoint Strategy}: Evaluation after each training epoch
    \item \textbf{Performance Tracking}: Weights \& Biases for experiment monitoring
\end{itemize}

\subsection{Controlled Variables and Experimental Validity}

\begin{itemize}
    \item \textbf{Hardware Consistency}: All experiments on identical A100-80GB GPUs
    \item \textbf{Software Environment}: Standardized Modal containers with fixed library versions
    \item \textbf{Data Consistency}: Identical training and evaluation data across all configurations
    \item \textbf{Random Seed Control}: Fixed seed for reproducible initialization and data shuffling
    \item \textbf{Hyperparameter Documentation}: Complete parameter logging for replication
\end{itemize}

%----------------------------------------------------------------------------------------